```{r setup, include=FALSE}
library(here)
library(scales)
library(tidyverse)
library(ggplot2)
library(modelr)

theme_set(theme_bw())

knitr::opts_chunk$set(echo = TRUE)
```

```{r Load_data}
trips_per_day <- read_tsv('trips_per_day.tsv')

#Manipulate the dataframe to include: weekends
trips_per_day <- trips_per_day |> mutate(
    day_name = weekdays(ymd),
)
trips_per_day$is_weekend <- ifelse(trips_per_day$day_name %in% c("Saturday", "Sunday"), 1, 0)

trips_per_day <- trips_per_day |> mutate(
    avg_temp = tmax - tmin,
    avg_prcp = mean(prcp)
)

trips_per_day <- trips_per_day |> mutate(
    month = month(ymd),
    season = case_when(
      month %in% c(12, 1, 2)  ~ "Winter",
      month %in% c(3, 4, 5)   ~ "Spring",
      month %in% c(6, 7, 8)   ~ "Summer",
      month %in% c(9, 10, 11) ~ "Fall"
    )
)


view(trips_per_day)



```

```{r setup the data}

set.seed(42)

# Train and validation set
num_days <- nrow(trips_per_day)
frac_train_validation <- 0.9
num_train_validation <- floor(num_days * frac_train_validation)

# randomly sample rows for the training & validation sets
ndx <- sample(1:num_days, num_train_validation, replace=F)
trips_per_day_train_validation <- trips_per_day[ndx, ]

# Trips for test
trips_per_day_test <- trips_per_day[-ndx, ]


# split the train and validation data
set.seed(42)
num_days_overall_train_validation <- nrow(trips_per_day_train_validation)
frac_train <- 0.6 
num_train <- floor(num_days_overall_train_validation * frac_train)
ndx_trian <- sample(1:num_days_overall_train_validation, num_train, replace = F)
trips_per_day_train <- trips_per_day_train_validation[ndx_trian, ]

# Trips for validation: 
trips_per_day_validate <- trips_per_day_train_validation [-ndx_trian, ]

```

```{r model}

# fit a model for each polynomial degree
K <- 1:8
train_err <- c()
validate_err <- c()
for (k in K) {
  
    # fit on the training data
    model <- lm(num_trips ~ poly(tmin, k, raw = T), data=trips_per_day_train)
    
    # evaluate on the training data
    train_err[k] <- sqrt(mean((predict(model, trips_per_day_train) - trips_per_day_train$num_trips)^2))

    # evaluate on the validate data
    validate_err[k] <- sqrt(mean((predict(model, trips_per_day_validate) - trips_per_day_validate$num_trips)^2))
}

```


```{r model plot}

plot_data <- data.frame(K, train_err, validate_err) %>%
  gather("split", "error", -K)

ggplot(plot_data, aes(x=K, y=error, color=split)) +
  geom_line() +
  scale_x_continuous(breaks=K) +
  xlab('Polynomial Degree') +
  ylab('RMSE')

```



```{r model prediction vs actual}


model <- lm(num_trips ~ poly(tmin, 5, raw = T), data = trips_per_day_train)

trips_per_day_train <- trips_per_day_train %>%
  add_predictions(model) %>%
  mutate(split = "train")
trips_per_day_validate <- trips_per_day_validate %>%
  add_predictions(model) %>%
  mutate(split = "validate")
plot_data <- bind_rows(trips_per_day_train, trips_per_day_validate)

ggplot(plot_data, aes(x = tmin, y = num_trips)) +
  geom_point(aes(color = split)) +
  geom_line(aes(y = pred)) +
  xlab('Minimum temperature') +
  ylab('Daily trips') +
  scale_y_continuous()
```


# New Models 


# Model with everything 

```{r model with avg temp and snow and is_weekend}

train_err <- c()
validate_err <- c()

view(trips_per_day_train)
    model_with_everything <- lm(num_trips ~ avg_temp +is_weekend + prcp + snow + snwd, data=trips_per_day_train)
    
    # evaluate on the training data
    train_err <- sqrt(mean((predict(model_with_everything, trips_per_day_train) - trips_per_day_train$num_trips)^2))

    # evaluate on the validate data
    validate_err <- sqrt(mean((predict(model_with_everything, trips_per_day_validate) - trips_per_day_validate$num_trips)^2))

    train_err # Gives the RMSE for the train # about  7997.772
    validate_err # Gives the RMSE for the train # about 7723.385

    summary(model_with_everything)   
```


# Model with rain
```{r model with rain}

train_err <- c()
validate_err <- c()

    model_with_rain <- lm(num_trips ~ (tmin + prcp), data=trips_per_day_train)
    
    # evaluate on the training data
    train_err <- sqrt(mean((predict(model_with_rain, trips_per_day_train) - trips_per_day_train$num_trips)^2))

    # evaluate on the validate data
    validate_err <- sqrt(mean((predict(model_with_rain, trips_per_day_validate) - trips_per_day_validate$num_trips)^2))

    train_err # Gives the RMSE for the train # about 5141 
    validate_err # Gives the RMSE for the train # about 4920 

    
```

# Model with rain and snow


```{r model with rain and snow}

train_err <- c()
validate_err <- c()

    model_with_rain_snow <- lm(num_trips ~ (tmin + prcp + snow), data=trips_per_day_train)
    
    # evaluate on the training data
    train_err <- sqrt(mean((predict(model_with_rain_snow, trips_per_day_train) - trips_per_day_train$num_trips)^2))

    # evaluate on the validate data
    validate_err <- sqrt(mean((predict(model_with_rain_snow, trips_per_day_validate) - trips_per_day_validate$num_trips)^2))

    train_err # Gives the RMSE for the train # about 5141 
    validate_err # Gives the RMSE for the train # about 4920 

    # Approximately no difference between it and just with rain 

    
```


# model with rain and snow and their interaction
```{r model with rain and snow and their interaction}

train_err <- c()
validate_err <- c()

    model_with_rain_snow_theirinteraction <- lm(num_trips ~ (tmin + prcp * snow), data=trips_per_day_train)
    
    # evaluate on the training data
    train_err <- sqrt(mean((predict(model_with_rain_snow_theirinteraction, trips_per_day_train) - trips_per_day_train$num_trips)^2))

    # evaluate on the validate data
    validate_err <- sqrt(mean((predict(model_with_rain_snow_theirinteraction, trips_per_day_validate) - trips_per_day_validate$num_trips)^2))

    train_err # Gives the RMSE for the train # about 5093 
    validate_err # Gives the RMSE for the train # about 4898 

    # Approximately no difference between it and just with rain 

    
```



# model with rain and snow and their interaction and their polynomial


```{r model}

# fit a model for each polynomial degree
K <- 1:8
train_err <- c()
validate_err <- c()
for (k in K) {
  
    # fit on the training data
    model_with_rain_snow_theirinteraction_with_poly <- lm(num_trips ~ poly((tmin + prcp * snow), k, raw = T), data=trips_per_day_train)
    
    # evaluate on the training data
    train_err[k] <- sqrt(mean((predict(model_with_rain_snow_theirinteraction_with_poly, trips_per_day_train) - trips_per_day_train$num_trips)^2))

    # evaluate on the validate data
    validate_err[k] <- sqrt(mean((predict(model_with_rain_snow_theirinteraction_with_poly, trips_per_day_validate) - trips_per_day_validate$num_trips)^2))
}

train_err # there are higher than the rain and snow and their interaction Not helpul

validate_err

```

# model with  rain and snow and their interaction & weekend effect & tmax

```{r model with rain and snow and their interaction & weekend effect &  tmax}

train_err <- c()
validate_err <- c()

view(trips_per_day_train)
    model_with_rain_snow_theirinter_avg_temp_weekend <- lm(num_trips ~ (tmax + is_weekend + prcp * snow), data=trips_per_day_train)
    
    # evaluate on the training data
    train_err <- sqrt(mean((predict(model_with_rain_snow_theirinter_avg_temp_weekend, trips_per_day_train) - trips_per_day_train$num_trips)^2))

    # evaluate on the validate data
    validate_err <- sqrt(mean((predict(model_with_rain_snow_theirinter_avg_temp_weekend, trips_per_day_validate) - trips_per_day_validate$num_trips)^2))

    train_err # 4143.062
    validate_err # 3938.039

    summary(model_with_rain_snow_theirinter_avg_temp_weekend)
    
```


# model with tmax interacting with snow depth and prcp interacting with snow  & weekend effect

```{r  model with tmax interacting with snow depth and prcp}

train_err <- c()
validate_err <- c()

view(trips_per_day_train)
    model_with_rain_snow_theirinter_avg_temp_weekend <- lm(num_trips ~ (tmax * snwd + is_weekend + prcp * snow), data=trips_per_day_train)
    
    # evaluate on the training data
    train_err <- sqrt(mean((predict(model_with_rain_snow_theirinter_avg_temp_weekend, trips_per_day_train) - trips_per_day_train$num_trips)^2))

    # evaluate on the validate data
    validate_err <- sqrt(mean((predict(model_with_rain_snow_theirinter_avg_temp_weekend, trips_per_day_validate) - trips_per_day_validate$num_trips)^2))

    train_err # 4051.437
    validate_err # 3810.801
```




# model with tmax interacting with snow depth and prcp interacting with snow  & weekend effect

```{r  model with tmax interacting with snow depth and prcp}

train_err <- c()
validate_err <- c()

view(trips_per_day_train)
    model_with_rain_snow_theirinter_temp_snwd_weekend_season <- lm(num_trips ~ (tmax * snwd + is_weekend * season + prcp * snow), data=trips_per_day_train)
    
    # evaluate on the training data
    train_err <- sqrt(mean((predict(model_with_rain_snow_theirinter_temp_snwd_weekend_season, trips_per_day_train) - trips_per_day_train$num_trips)^2))

    # evaluate on the validate data
    validate_err <- sqrt(mean((predict(model_with_rain_snow_theirinter_temp_snwd_weekend_season, trips_per_day_validate) - trips_per_day_validate$num_trips)^2))

    train_err # 3620
    validate_err # 4108.801

    #summary(model_with_rain_snow_theirinter_temp_snwd_weekend_season)
```






# Let's see which degree is the best for every variable: 

``` {r model check variables}

K <- 1:8
train_err <- c()
validate_err <- c()
for (k in K) {
  
    # fit on the training data
    model <- lm(num_trips ~ poly(snwd, k, raw = T), data=trips_per_day_train)
    
    # evaluate on the training data
    train_err[k] <- sqrt(mean((predict(model, trips_per_day_train) - trips_per_day_train$num_trips)^2))

    # evaluate on the validate data
    validate_err[k] <- sqrt(mean((predict(model, trips_per_day_validate) - trips_per_day_validate$num_trips)^2))
}


```


``` {r model check the best degree for the variables}
plot_data <- data.frame(K, train_err, validate_err) %>%
  gather("split", "error", -K)

ggplot(plot_data, aes(x=K, y=error, color=split)) +
  geom_line() +
  scale_x_continuous(breaks=K) +
  xlab('Polynomial Degree') +
  ylab('RMSE')
```



# My best model so far 

```{r  best model}

train_err <- c()
validate_err <- c()

view(trips_per_day_train)
    model_with_rain_snow_theirinter_temp_snwd_weekend_season <- lm(num_trips ~ (tmax * snwd + is_weekend * season + prcp * snow), data=trips_per_day_train)
    
    # evaluate on the training data
    train_err <- sqrt(mean((predict(model_with_rain_snow_theirinter_temp_snwd_weekend_season, trips_per_day_train) - trips_per_day_train$num_trips)^2))

    # evaluate on the validate data
    validate_err <- sqrt(mean((predict(model_with_rain_snow_theirinter_temp_snwd_weekend_season, trips_per_day_validate) - trips_per_day_validate$num_trips)^2))

    train_err # 3620
    validate_err # 4108.801

    #summary(model_with_rain_snow_theirinter_temp_snwd_weekend_season)
```


# My Plot for the best model graph 1 

```{r plot best model graph 1}

trips_per_day_train <- trips_per_day_train %>%
  add_predictions(model_with_rain_snow_theirinter_temp_snwd_weekend_season) %>%
  mutate(split = "train")

trips_per_day_validate <- trips_per_day_validate %>%
  add_predictions(model_with_rain_snow_theirinter_temp_snwd_weekend_season) %>%
  mutate(split = "validate")

plot_data <- bind_rows(trips_per_day_train, trips_per_day_validate)

view(plot_data)

ggplot(plot_data, aes(x = date)) +
  geom_point(aes(y = num_trips, color = split)) +
  geom_line(aes(y = pred)) +
  xlab('Date ') +
  ylab('Daily trips') +
  scale_y_continuous()

```


# My Plot for the best model graph 2 

```{r plot best model graph 2}

view(plot_data)

ggplot(plot_data) +
  geom_point(aes(x = pred, y = num_trips)) +
  geom_abline(color ="blue") +
  xlab('Prediction') +
  ylab('Daily trips') +
  scale_y_continuous()

```

# save my best model

```{r save my best model}
save(model_with_rain_snow_theirinter_temp_snwd_weekend_season, file = "Bestmodel.RData")
```








